The impacts of park access on health outcomes: a spatial comprehensive approach

### 2018-12-12 manuscript.Rnw
Add data summary table

\maketitle
<<setup, cache=FALSE, echo=FALSE, message=FALSE>>=
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
knitr::opts_chunk$set(cache=TRUE)

@@

@@

# identify complete cases
complete_tracts_index <- tracts_sf %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  complete.cases()
tracts <- tracts[complete_tracts_index, ]
@

@@

@@

\subsection{Park Choice Logsum}
In the absence of a park choice survey, we estimated likely values of the $\beta$
coefficients by iteratively searching for the values which produced the highest
model likelihood in *which model did we use to estimate likelihood?*. The search
was constrained by two assertions. First, we required that the coefficient on
size be positive and that on distance negative; all else equal, people will
prefer to use larger parks that are nearer to their residence. Second, we visually
inspected the resulting accessibility scores to ensure that the scores produced a
reasonably varied pattern of access throughout the metropolitan region.

$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
We wrote a function to calculate this value for an arbitrary distance matrix, size
and tweets vector, and coefficients.

@@

It may in some cases be more efficient to estimate a restricted spatial autoregressive
model, that controls for only correlation or dependence. \citet{tra_autoregression_paper}
illustrate a decision algorithm --- proposed by \citet{Florax2003} --- which uses
a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
if the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used.

>>>>

\maketitle
<<setup, cache=FALSE, echo=FALSE, message=FALSE>>=
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
opts_chunk$set(cache=TRUE)

@@

Note that this formulation is an extension of the gravity-based accessibility 
statistics, in the same way that the gravity model is a specific case of 
a destination-choice model \citep{Daly1982}. The extension

@@

# identify complete cases
complete_tracts_index <- tracts_sf %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  complete.cases()
tracts <- tracts[complete_tracts_index, ]

tracts@data <-  tbl_df(tracts@data) %>%
  dplyr::transmute(
    county_fips = substr(GEOID, 3, 5),
    borough = case_when(
      county_fips == "081" ~ "Queens",
      county_fips == "047" ~ "Brooklyn",
      county_fips == "061" ~ "Manhattan",
      county_fips == "005" ~ "Bronx",
      county_fips == "085" ~ "Staten Island",
      TRUE ~ as.character(NA)
    ),
    obesity = OBESITY,
    physact = 100 - Phys_Act,
    log_obesity = log(OBESITY),
    log_physact = log(100 - Phys_Act),
    density = Pop_Density,
    fulltime = FulltimeWork, college = CollegeDeg, 
    single = Single_Percent,
    black = PctBlack, asian = PctAsian, hispanic = PctHispanic,
    other = PctNative + PctPacific,
    Pct0to17, Pct18to29, Pct65plus, 
    Income1, Income2, Income3, Income4, Income5, Income6, 
    Income7, Income8, Income9, Income10
  )

tracts_sf <- st_as_sf(tracts)
@

@@

\begin{table*}
  \caption{Summary Statistics of Tract Data}
  \label{tab:parkbenefits}
<<data_summary, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
our_summary <- list(
  "obesity"     = tab_summary(tracts$obesity),
  "physact"     = tab_summary(tracts$physact),
  "pop_density" = tab_summary(tracts$density),
  "fulltime"    = tab_summary(tracts$fulltime),
  "college"     = tab_summary(tracts$college)
)

summary_table(tracts@data, our_summary) %>%
  kable("latex", booktabs = T, escape = FALSE) %>%
  group_rows("Obese: percent of tract which is obese", 1, 4) %>%
  group_rows("Physical Activity: Percent of tract reporting physical activity", 5, 8) %>%
  group_rows("Population Density: individuals per square mile", 9, 12)  %>%
  group_rows("Full Time: percent of adults who work full time", 13, 16) %>%
  group_rows("College: percent of adults with a college degree", 17, 20)
@
\end{table*}

@@

In the absence of a park choice survey, we estimated likely values of the $\beta$
coefficients by iteratively searching for the values which produced the highest
model likelihood in *which model did we use to estimate likelihood?*. The search
was constrained by two assertions. First, we required that the coefficient on
size be positive and that on distance negative; all else equal, people will
prefer to use larger parks that are nearer to their residence. Second, we visually
inspected the resulting accessibility scores to ensure that the scores produced a
reasonably varied pattern of access throughout the metropolitan region.

<<f_logsums>>=
#' Calculate destination choice logsums from a distance matrix
#' 
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients. 
#'   If we submit two variables the tweets are ignored.
#'   
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a 
#' 
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-15,.00001, .001)){

  # A is n x p
  a <- betas[1] * d

  # B is p x 1
  b <- betas[2] * sizes

  if(!is.null(tweets)) 
    b <- b + betas[3] * tweets

  # calculate observed utility by adding the weighted park-level attributes 
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(a, 2, b, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))

}
@

@@

It may in some cases be more efficient to estimate a restricted spatial autoregressive
model, that controls for only correlation or only dependence. \citet{tra_autoregression_paper}
illustrate a decision algorithm --- proposed by \citet{Florax2003} --- which uses
a likelihood ratio test of the SDM against a restricted spatial error model (SEM):
if the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used.
